---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r warning=FALSE}    
library(ggplot2)
  
# Load data from csv file
activity_data <- read.csv("activity.csv")
#Strip out NAs
noNA_data <- activity_data[complete.cases(activity_data[,]),]
#convert to date
noNA_data$date <- as.Date(as.character(noNA_data$date),"%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r}
agg_noNA_data <- aggregate(noNA_data$steps, by=list(date=noNA_data$date), FUN=sum)
names(agg_noNA_data)[2]<-"steps"

#plot
qplot(date,steps,data=agg_noNA_data,geom = "bar",stat="identity",main="Total Steps Per Day")

#mean & median
noNA_mean_median <- aggregate(noNA_data$steps, by=list(date=noNA_data$date), FUN=mean)
names(noNA_mean_median)[2]<-"mean_steps"
noNA_median_steps <- aggregate(noNA_data$steps, by=list(date=noNA_data$date), FUN=median)
noNA_mean_median <- merge(noNA_mean_median, noNA_median_steps, by = c("date"), all.x = TRUE)
names(noNA_mean_median)[3]<-"median_steps"
#Show values
noNA_mean_median
```



## What is the average daily activity pattern?
```{r}
intervalAgg <- aggregate(noNA_data$steps, by=list(date=noNA_data$interval), FUN=mean)
names(intervalAgg)[1]<-"interval"
names(intervalAgg)[2]<-"steps"
ts <- ts(intervalAgg$steps,frequency=24*12)
plot(intervalAgg$interval,ts,type="l",xlab="Interval",ylab="Mean Steps",main="Average Number Of Steps Taken")

maxSteps <- max(intervalAgg$steps)
maxstepinterval <- intervalAgg[which(intervalAgg$steps == maxSteps),1] 
```
The 5-minute interval that contains `r maxSteps`, which is the maximum number of steps, is interval `r maxstepinterval`

## Imputing missing values

My strategy is to calculate the mean for each 5 minute interval and apply that to NA values.

```{r}
# Load data 
all_activity_data <- activity_data

#use mean calulated in intervalAgg to replace NAs
modifiedMeans <- merge(all_activity_data, intervalAgg, by = c("interval"), all.x = TRUE)
modifiedMeans <- transform(modifiedMeans, steps = ifelse(is.na(steps.x), steps.y, steps.x), steps.x = NULL, steps.y = NULL)

#convert to date
modifiedMeans$date <- as.Date(as.character(modifiedMeans$date),"%Y-%m-%d")

#plot
qplot(date,steps,data=modifiedMeans,geom = "bar",stat="identity",main="Total number of steps taken each day after missing values were imputed")

#mean & median
mod_mean <- aggregate(modifiedMeans$steps, by=list(date=modifiedMeans$date), FUN=mean)
names(mod_mean)[2]<-"mean_steps"
mod_median <- aggregate(modifiedMeans$steps, by=list(date=modifiedMeans$date), FUN=median)
names(mod_median)[2]<-"median_steps"

mod_mean_median <- aggregate(modifiedMeans$steps, by=list(date=modifiedMeans$date), FUN=mean)
names(mod_mean_median)[2]<-"mean_steps"
mod_median_steps <- aggregate(modifiedMeans$steps, by=list(date=modifiedMeans$date), FUN=median)
mod_mean_median <- merge(mod_mean_median, mod_median_steps, by = c("date"), all.x = TRUE)
names(mod_mean_median)[3]<-"median_steps"

#Show values
mod_mean_median
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
#Work out whether date is weekday or weekend 
modifiedMeans$weekdays <- weekdays(modifiedMeans$date)
modifiedMeans <- transform(modifiedMeans, weekdays = ifelse(weekdays %in% c("Saturday","Sunday"), "weekend", "weekday"))
modifiedMeans$weekdays <- factor(modifiedMeans$weekdays)

tsWeekday <- ts(modifiedMeans[which(modifiedMeans$weekdays == "weekday"),3],frequency=24*12)
tsWeekend <- ts(modifiedMeans[which(modifiedMeans$weekdays == "weekend"),3],frequency=24*12)

par(mfrow = c(2, 1),mar = c(2,4,1,1))
plot(modifiedMeans[which(modifiedMeans$weekdays == "weekday"),1],tsWeekday,type="l",xlab="Interval",ylab="Mean Steps",main="Weekday",col=3)
par(mar = c(2,4,1,1))
plot(modifiedMeans[which(modifiedMeans$weekdays == "weekend"),1],tsWeekend,type="l",xlab="Interval",ylab="Mean Steps",main="Weekend",col=2)


```